---
title: "Western Sydney Strategic Plan –   \n Threatened Species Habitat Mapping"
author: "Ascelin Gordon  \n Vira Koshkina  \n RMIT University"
# date: "`r Sys.Date()`"
output:
   pdf_document:
      fig_caption: true
      number_sections: true
classoption: landscape

---
\newpage

\tableofcontents

\newpage
## Predictor layers used in the first step of the analysis
ce_radann - Annual Mean Radiation
ce_radhp - Highest Period Radiation
ce_radlp - Lowest Period Radiation
ct_frostdays_lt2deg_1 - Number of days/annum with minimum temperature less than 2 degrees ct_tempann_1 - Mean annual temperature
ct_tempseas_1 - Temperature of Seasonality: Coefficient of Variation
cw_precipseas - Precipitation of Seasonality: Coefficient of Variation
DEM_c - The altitude of a cell above sea level (25m DEM of Hunter Valley)
dl_lat_grid - Latitude (surrogate for location, disperal, isolation)
dl_strmdstge2 - Euclidean distance to 2nd order streams and above
dl_strmdstge4 - Euclidean distance to 4th order streams and above
dl_strmdstge6 - Euclidean distance to 6th order streams and above
gp_k_fillspl - filtered potassium (K), gaps filled in using geograhically weighted regession model and spline function
lf_aspect_tr - Beer's Apsect- transformation of aspect to a continuous scaled variable. Changed for the southern hemisphere by setting maximum value (2) to SE slopes (coolest) and minimum (0) to NW slopes (warmest).
lf_cti - Compound topographic index
lf_rough0100 - Neighbourhood topographical roughness based on the standard deviation of elevation in a circular 100 m neighbourhood. lf_tpi0120 - Topographic position index using neighbourhood of 120m radius
sfc_aut_b_95 - Seasonal Fractional Cover representing proportions of green (g), bare (b), and non-green or senescent (b). Derived from Landsat imagery over the period of 1998 to 2012, for the season of autumn and for 95th percentile.
soil – Soil type
vegitation_pct_b – Native vegetation (PCT)

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r prep, include=FALSE}
require(rgdal)
require(raster)
require(maptools)
require(dismo)
library(sp)
library(ENMeval)
library(caTools)
dyn.load('/Library/Java/JavaVirtualMachines/jdk1.8.0_181.jdk/Contents/Home/jre/lib/server/libjvm.dylib')
library(rJava)


source('utilities.R')


species_list =c("Litoria_aurea","Grevillea_parviflora_subsp_parviflora", "Persoonia_nutans","Phascolarctos_cinereus"  )

data_folder = "Data/Cumberland"
presence_folder = "Data/Presence"
bias_folder = "Data/Bias"


#read all the covariate tif files from covariate folder
cumberland <-  readOGR("Data/IBRA_cumberland/Cumberland_IBRA_subregion.shp")
bias = readOGR("Data/bias/presence_layer_buffer.shp")

files=list.files(data_folder , pattern = "*.tif$", full.names=TRUE)
no_cov=length(files) #number of covariates
predictors <- stack(files)




### add zeros instead of NA in the vegetation layer
predictors <- dropLayer(predictors, 'vegitation_pct_b_c')
vegetation=raster(paste(data_folder, '/', 'vegitation_pct_b_c.tif', sep=''))
vegetation[is.na(vegetation[])] <- 0
plot(vegetation)
predictors <- addLayer(predictors, vegetation)

background_sample_file = "Data/background_sample.rda"


load_sample_file = function(background_sample_file, bias){
    ppi=300
    if(!file.exists(background_sample_file)) {
        background_sample = mask(predictors, bias)
        plot(background_sample)
        save(background_sample,  file=background_sample_file)
        # writeRaster(background_sample, background_sample_file, format = "GTiff", overwrite=TRUE)
        
    }else{
        load(background_sample_file)
    }
    return(background_sample)
}    

get_bkgr_train_test = function(backgr_train_file, background_sample){
    if(!file.exists(backgr_train_file)){
        #### select beckground points - randomly selecting points and save them for later
        backgr <- randomPoints(background_sample, 1000)
        group <- kfold(backgr, 5)
        backgr_train <- backgr[group != 1, ]
        backgr_test <- backgr[group == 1, ]
        #
        save(backgr_train, backgr_test, file=backgr_train_file)
        
        print(paste("Testing and training points for bakground have been randomly selected and saved in the file: ", backgr_train_file))
        
    }else{
        load(backgr_train_file)
        print(paste("Testing and training points for background have been loaded from the file: ", backgr_train_file))
    }
    return(c(backgr_train, backgr_test))
}



```

```{r echo = FALSE, results ='asis'}

for (species in species_list){
    
    # create a folder for output
    dir.create(paste('Outcomes/', species, sep=''), showWarnings = FALSE)
    
    pres_train_file = paste('Outcomes_Cumberland/', species, '/',"pres_test_train_", species, ".rda" , sep="")
    backgr_train_file = paste("bias_backgr_test_train.rda")
    

    
    
    presence_table <- read.table(paste(presence_folder,"/", species, ".csv", sep="" ),  header=TRUE,  sep=",")
    presences_all = SpatialPoints( presence_table[, c("decimalLon","decimalLat")])
    proj4string(presences_all)=cumberland@proj4string
    
    # only select points in the study area
    presences = presences_all[cumberland,]
    
    # check if training and testing points have already been selected
    if(!file.exists(pres_train_file)){
        ##### select 20% of the data for testing - randomly selecting points
        group <- kfold(presences, 5)
        pres_train <- presences[group != 1, ]
        pres_test <- presences[group == 1, ]
        
        save(pres_train, pres_test, file=pres_train_file)
        
        print(paste("Testing and training points for", species, "(", species, ") have been randomly selected and saved in the file: ", pres_train_file))
        
        
    }else{
        load(pres_train_file)
        print(paste("Testing and training points for", species, "(", species, ") have been loaded from the file: ", pres_train_file))
    }
    
    png(filename =paste('Outcomes_Cumberland/', species, '/', species, "_presence_background_points.png", sep=""), width = 6*ppi, height = 3*ppi)
    # plot presence and bacground points
    plot(!is.na(predictors[[1]]), col=c('white', 'light grey'), legend=FALSE)
    points(backgr_train, pch='-', cex=1.5, col='yellow')
    points(backgr_test, pch='-',  cex=1.5, col='black')
    points(pres_train, pch= '+',cex=1.5, col='green')
    points(pres_test, pch='+',cex=1.5, col='blue')
    
    legend("topright", legend=c("background - train","background - test","presence - train","presence - test"), pch=c('-','-','+','+'), col = c('yellow', 'black', 'green','blue'),cex=2.5)
    dev.off()

    
    # ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
    # RUN MAXENT - all covariates --------------------------------------------------
    # ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
    
    factors = c('vegitation_pct_b_c', 'soil_c')
    
    model_name = "select_variables"
    outcome_folder = 'Outcomes_Cumberland/'
 
     #  
    model = run_maxent_reduce_variables (predictors, factors, species,  model_name, pres_train, pres_test, backgr_train, backgr_test, outcome_folder)
    
    # ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
    # RUN MAXENT - bias layers --------------------------------------------------
    # ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
    
    species_bias = paste (bias_folder, species, sep="")
    
    bias_list=c( common_bias, species_bias)
    
    
    # ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
    # RUN MAXENT - vegetation --------------------------------------------------
    # ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
    
    
    }
```


```{r echo = FALSE, results ='asis'}

species_list=c("A", "B","C")

for(species in species_list){
  cat('\n##', species, '\n')
	cat('\\pagebreak')
	cat('\n')
}

```


